# Part B. Reshaping data.

The file `rhpi.csv` contains data from the International House Price Database from the Federal Reserve Bank of Dallas. It records the real housing price index. The raw data is available [here](https://www.dallasfed.org/institute/houseprice/~/media/documents/institute/houseprice/hp1701.xlsx): 

Open the `.csv` file into `R`.

```{r}
rhpi <- read.csv("rhpi.csv", stringsAsFactors = FALSE)
```

1. Use `head` to explore the first few rows of the dataset. Is it in long or wide format? Is it a `tidy` dataset? Why or why not? (10 points)

```{r}
head(rhpi)
```

__It is in wide format. It is not a tidy dataset because multiple columns correspond to the same variable.__

2.  You may find out that the first column name is empty. You can update the first column name to `country` by using the `colnames` function (5 points)

```{r}
colnames(rhpi)[1] <- "country"
```

3. Reshape the dataset so that it becomes _tidy_. Note that in most cases this means reshaping from wide to long. You can use the `melt` function from the `reshape` package or other functions. (15 points).

```{r}
library(reshape2)
rhpi.long <- melt(rhpi, id.vars="country", value.name="index")
head(rhpi.long)
```

4. Filter the data frame so that it only contains the data that corresponds to the US. Then, produce a line plot that tracks the housing price over time. What do you find? (10 points)

```{r}
us <- rhpi.long[rhpi.long$country=="US",]
plot(us$variable, us$index, type="l")

library(lubridate)
us$variable <- gsub('X|Q', '', as.character(us$variable))
us$q <- parse_date_time(us$variable, orders = "Yq")
plot(us$q, us$index, type="l", xlab="Year/Quarter", ylab="Housing Price Index")
```

_We can clearly see the housing bubble and then the financial crisis after 2008._

5. Write a function that takes a country name as input and returns the average housing price in the last 5 years. Then, show that it works by running it for 3 different countries. (10 points)

```{r}
rhpi.long$year <- substr(as.character(rhpi.long$variable), 2, 5)

avg_hp <- function(country){
  df <- rhpi.long[rhpi.long$country==country & rhpi.long$year > 2012,]
  return(mean(df$index))
}

avg_hp("Australia")
avg_hp("Belgium")
avg_hp("Canada")

```




